---
title: "Apology Data Analysis"
output: 
  pdf_document:
    toc: true
date: "2024-02"
---

## Import Packages
```{r setup, message=FALSE, warning=FALSE}
library(readxl)
library(ggplot2)
library(tidyverse)
library(dplyr)
```

# Task 1 - Answer a Research Question
```{r read-in data, message=FALSE, warning=FALSE}
# First, I load the given dataset in my RStudio. 
# I change the file name to data for easier access
key <- read_excel("data.xlsx", sheet = "Key - variable meaning") 
fulldata <- read_excel("data.xlsx", sheet = "Data")

# Next, I do some preliminary data cleaning to drop unreliable participants
df <- fulldata %>% 
  filter(Finished == TRUE) %>% 
  filter(passedattn ==  "yes") %>% 
  # there is one comment saying "don't use this data, with a age of 149, so I filter this out using age as a criterion
  filter(age < 100)
```

```{r , message=FALSE, warning=FALSE}

```
## 1. Whether people care about getting a return apology after being the first to apologize
```{r , message=FALSE, warning=FALSE}
compare <- df %>%
  select(feelings_youalone, feelings_bothyoufirst) %>%
  rename('Get no return apology' = feelings_youalone, 
         'Get return apology' = feelings_bothyoufirst) %>%
  pivot_longer(cols = everything(), names_to = "scenario", values_to = "score")

ggplot(compare, aes(x = scenario, y = score, fill = scenario)) +
  geom_boxplot() +
  labs(title = "Whether people care about receiving an apology",
       subtitle = "after being the first to apologize",
       x = "Scenario",
       y = "Feeling Score",
       fill = "Scenario Type") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1")

```
create a bar plot People get a return apology after being the first to apologize feelings_both you first
compare with people apologize and did not get apologies back 
## 2. Whether this varies as function of individual differences in “initiator type”
```{r , message=FALSE, warning=FALSE}

compare_type <- df %>%
  select(feelings_youalone, feelings_bothyoufirst, initiator_type) %>%
  rename('Get no return apology' = feelings_youalone, 
         'Get return apology' = feelings_bothyoufirst) %>%
  pivot_longer(cols = c('Get no return apology', 'Get return apology'), names_to = "scenario", values_to = "score")

# Now create the plot with facets
ggplot(compare_type, aes(x = scenario, y = score, fill = scenario)) +
  geom_boxplot() +
  labs(title = "Whether people care about receiving an apology",
       subtitle = "after being the first to apologize",
       x = "Scenario",
       y = "Feeling Score",
       fill = "Scenario Type") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))+
  scale_fill_brewer(palette = "Set1") +
  facet_wrap(~initiator_type)

```
## 3. whether a return apology is simply viewed as a form of forgiveness
```{r , message=FALSE, warning=FALSE}

```
# Task 2 - Conduct additional analyses
## A. Produce a bar graph
```{r plot feelings variable, message=FALSE, warning=FALSE}
# Calculate means and standard errors
feelings <- df %>%
  select("feelings_youalone", "feelings_bothyoufirst", "feelings_themalone",
         "feelings_boththemfirst","feelings_neither","feelings_youaloneforgiven")
feelings_means <- colMeans(feelings)
feelings_se <- sapply(feelings, function(x) sd(x) / sqrt(length(x)))
plot_feelings <- data.frame(scenario = names(feelings_means),
                            mean = feelings_means,
                            se = feelings_se) 

# Create the bar graph with ggplot2
ggplot(plot_feelings, aes(x = reorder(scenario, -mean), 
                          y = mean, fill = scenario)) +
  geom_bar(stat = "identity", 
           position = "dodge") +
  geom_errorbar(aes(ymin = mean - se, 
                    ymax = mean + se),
                width = .2,                   
                position = position_dodge(.9)) +
  labs(x = "Scenario", 
       y = "Average of Feelings", 
       title = "Average of Feelings for Each Scenario",    
       fill = "Scenario Type") +
  scale_fill_manual(values=c("#ff6db6","#F46D43","#490092",
                             "#006ddb","#24ff24","#ffff6d"))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
    panel.background = element_rect(fill = "white"),   
    panel.grid.major.y = element_line(colour = "lightgray", linetype = 1, size = .5),
    panel.grid.major.x = element_line(colour = "lightgray", linetype = 1, size = .5),
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank())
```
## B. Conduct ANOVA
```{r ANOVA, message=FALSE, warning=FALSE}
# Perform one-way ANOVA
anova_data <- feelings %>% pivot_longer(cols = everything(), names_to = "scenario", values_to = "score")
anova_results <- aov(feelings_youalone ~ feelings_bothyoufirst + feelings_themalone + feelings_boththemfirst + feelings_neither + feelings_youaloneforgiven, data = df)
summary(anova_results)
```

### Pairwise t-tests
```{r paired t-tests, message=FALSE, warning=FALSE}
# Exclude the first column 'feelings_youalone' for the pairwise comparisons
scenarios <- names(feelings)[-1] 

# Initialize an empty list to store test results and a vector for p-values
pairwise_tests <- list()
p_values <- numeric(length(scenarios))

# Perform pairwise t-tests for all pairs where 'feelings_youalone' is one group
for (scenario in scenarios) {
  test_result <- t.test(feelings[[scenario]], 
                        feelings[['feelings_youalone']], paired = TRUE)
  pairwise_tests[[scenario]] <- test_result
  scenario_index <- which(names(feelings) == scenario)
  # Adjusting index as 'feelings_youalone' is at index 1
  p_values[scenario_index - 1] <- test_result$p.value 
}

# Adjust for multiple testing using Bonferroni correction for multiple comparison 
p_adjusted <- p.adjust(p_values, method = "bonferroni")

# Combine results into a data frame
results <- data.frame(
  Scenario = scenarios,
  P_Value = p_values,
  P_Adjusted = p_adjusted,
  options(scipen=9999))
```
