---
title: "Apology Data Analysis"
output: 
  pdf_document:
    toc: true
header-includes:
  - \usepackage{booktabs}
date: "2024-02"
---

## Import Packages
```{r setup, message=FALSE, warning=FALSE}
library(readxl)
library(ggplot2)
library(tidyverse)
library(dplyr)
library(knitr)
```

# Task 1 - Answer a Research Question
```{r read-in data, message=FALSE, warning=FALSE}
# First, I load the given dataset in my RStudio. 
# I change the file name to data for easier access
key <- read_excel("data.xlsx", sheet = "Key - variable meanings") 
fulldata <- read_excel("data.xlsx", sheet = "Data")

# Next, I do some preliminary data cleaning to drop unreliable participants
df <- fulldata %>% 
  filter(Finished == TRUE) %>% 
  filter(passedattn ==  "yes") %>% 
  # there is one comment saying "don't use this data, with a age of 149, so I filter this out using age as a criterion
  filter(age < 100)
```

```{r , message=FALSE, warning=FALSE}

```
## 1. Whether people care about getting a return apology after being the first to apologize
```{r receiving apology, message=FALSE, warning=FALSE}
compare <- df %>%
  select(feelings_youalone, feelings_bothyoufirst) %>%
  rename('Get no return apology' = feelings_youalone, 
         'Get return apology' = feelings_bothyoufirst) %>%
  pivot_longer(cols = everything(), 
               names_to = "scenario", 
               values_to = "score")

ggplot(compare, aes(x = scenario, 
                    y = score, 
                    fill = scenario)) +
  geom_boxplot() +
  labs(title = "Whether people care about receiving an apology",
       subtitle = "after being the first to apologize",
       x = "Scenario",
       y = "Feeling Score",
       fill = "Scenario Type") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1")
```

To investigate whether people care about receiving an apology, I create a box plot to compare the feelings score distribution across two variables: feelings_youalone (in which people apologize first but receive no return apology) and feelings_bothyoufirst (in which people apologize first and receive other's apologies). I compare these two variables because the difference between these two scenarios is whether people get a return apology after they apologize first, and the difference of their scores could illustrate whether people care about get a return apology.\
From the box plot, I could tell that the general distribution of feelings when people get a return apology is much higher than the case of people get no return apology. The median of feelings get a return apology is much higher than people get no return apology (10 versus -20), and this also happens to the 25% quantile and 75% quantile bar. Thus, the conclusion is people actually cares very much about where they get a return apology when being the first to apologize.

## 2. Whether this varies as function of individual differences in “initiator type”
```{r initiator type, message=FALSE, warning=FALSE}

compare_type <- df %>%
  select(feelings_youalone, feelings_bothyoufirst, initiator_type) %>%
  rename('Get no return apology' = feelings_youalone, 
         'Get return apology' = feelings_bothyoufirst) %>%
  pivot_longer(cols = c('Get no return apology', 'Get return apology'), 
               names_to = "scenario", 
               values_to = "score")

# Now create the plot with facets
ggplot(compare_type, aes(x = scenario, 
                         y = score, 
                         fill = scenario)) +
  geom_boxplot() +
  labs(title = "Whether people care about receiving an apology by initiator type",
       subtitle = "(after being the first to apologize)",
       x = "Scenario",
       y = "Feeling Score",
       fill = "Scenario Type") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
        strip.background = element_rect(fill = "grey"),
        panel.grid.major.y = element_line(color = "grey", linetype = 1, size = .5),
        panel.grid.major.x = element_line(color = "grey", linetype = 1, size = .5),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.background = element_rect(fill = "white"))+
  scale_fill_brewer(palette = "Set1") +
  facet_wrap(~initiator_type)

```

If we investigate whether people care about getting a return apology by initiator types, we could get the same conclusion. For each initiator type, the feelings score distributions for the condition of getting a return apology is much higher than the condition of getting no return apology. 

## 3. whether a return apology is simply viewed as a form of forgiveness
```{r forgiveness, message=FALSE, warning=FALSE}

compare_forgive <- df %>%
  select(feelings_bothyoufirst, feelings_youaloneforgiven) %>%
  rename('Get return apology' = feelings_bothyoufirst, 
         'Get forgiveness' = feelings_youaloneforgiven) %>%
  pivot_longer(cols = everything(), 
               names_to = "scenario", 
               values_to = "score")

ggplot(compare_forgive, aes(x = scenario, 
                            y = score, 
                            fill = scenario)) +
  geom_boxplot() +
  labs(title = "Whether a return apology is viewed as a form of forgiveness",
       x = "Scenario",
       y = "Feeling Score",
       fill = "Scenario Type") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1")

```
In this case, I create a boxplot comparing the scenario of whether people get a return apology (feelings_bothyoufirst) or get a forgiveness (feelings_youaloneforgiven) to see if a return apology is simply viewed as a form of forgiveness. Since the distribution for the situation of get a return apology is much higher than the situation of get a forgiveness, we could tell that receiving a return apology is not the same as receiving a forgiveness. Receiving an actual return apology has a much higher score of feelings (median = 10), compared to get a forgiveness (median = -20). Thus, people do not just simply view a return apology as a form of forgiveness. 

## Some possible psychological explanations 
- In all three cases above, we could tell that feelings score of people getting a return apology is always higher whatever the initiator type or people get a forgiveness. This could due to several reasons:\
1. Mismatched expectations: People have inherent expectations of reciprocity and equity in social interactions. When we apologize, we might also expect the receiver to apologize as well to form a mutual apology. When they fail to do so, we might feel disappointed or injustice.\
2. Social norms: Similar to unmet expectations, in the case of non-reciprocated apology, we could feel being rejected by social norms or misunderstood, which could leads to social awkwardness or social isolated.\
3. Downwards self-esteem: Apologizing can make people feel vulnerable because we expose our feelings and lower our self-esteem. When we get a return apology, our self-esteem is protected, whereas hurt if we cannot get a return apology.\
- There could also be some potential factors and contexts influencing the results:\
1. Social status: I imagine if we could get return apologize to a person at a higher social positions than us, such as leaders or other respectful persons, whom we normally would not expect return apology from them, we might have a higher feelings score. First, since we do not usually expect their mutual apology, getting one would result in an experience of being respected and justice. Also, we could feel our self-esteem is being acknowledged when we get apology from people come from higher social hierarchic. 
2. Cultural norms: I think culture also plays as another factor could result in different feelings score here. People origin from different cultures might have different ways to express their apology or receive apology. The cultural discrepancy could leads to misunderstanding within senders and receivers. Even if senders accept a return apology, if the format is not applicable, their feeling score might be lower than normal.\ 
- Other Observations:
1.

# Task 2 - Conduct additional analyses
## A. Bar graph: average feelings across 6 scenarios
```{r plot feelings variable, message=FALSE, warning=FALSE}
# Calculate means and standard errors
feelings <- df %>%
  select("feelings_youalone", "feelings_bothyoufirst", "feelings_themalone",
         "feelings_boththemfirst","feelings_neither","feelings_youaloneforgiven")
feelings_means <- colMeans(feelings)
feelings_se <- sapply(feelings, function(x) sd(x) / sqrt(length(x)))
plot_feelings <- data.frame(scenario = names(feelings_means),
                            mean = feelings_means,
                            se = feelings_se) 

# Create the bar graph with ggplot2
ggplot(plot_feelings, aes(x = reorder(scenario, -mean), 
                          y = mean, fill = scenario)) +
  geom_bar(stat = "identity", 
           position = "dodge") +
  geom_errorbar(aes(ymin = mean - se, 
                    ymax = mean + se),
                width = .2,                   
                position = position_dodge(.9)) +
  labs(x = "Scenario", 
       y = "Average of Feelings", 
       title = "Average of Feelings for Each Scenario",    
       fill = "Scenario Type") +
  scale_fill_manual(values=c("#ff6db6","#F46D43","#490092",
                             "#006ddb","#24ff24","#ffff6d"))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
    panel.background = element_rect(fill = "white"),   
    panel.grid.major.y = element_line(color = "lightgray", linetype = 1, size = .5),
    panel.grid.major.x = element_line(color = "lightgray", linetype = 1, size = .5),
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank())
```
The average feelings score for each 6 scenario further validates our findings in previous part. People get better feelings score when they get mutual apology, and the score is even higher if others be the one to apologize first. In contrast, people feel the worst if they apologize first, but others do not apologize back.\

## B.1 Conduct ANOVA
```{r ANOVA, message=FALSE, warning=FALSE}
# Perform one-way ANOVA
anova_data <- feelings %>% pivot_longer(cols = everything(), 
                                        names_to = "scenario", 
                                        values_to = "score")
anova_results <- aov(score ~ scenario, data = anova_data)
summary_anova <- summary(anova_results)
anova_table <- data.frame(
  Term = rownames(summary_anova[[1]]),
  Df = summary_anova[[1]]$Df,
  Sum_Sq = summary_anova[[1]]$`Sum Sq`,
  Mean_Sq = summary_anova[[1]]$`Mean Sq`,
  F_value = summary_anova[[1]]$F,
  Pr_F = summary_anova[[1]]$`Pr(>F)`
)
kable(anova_table, format = "latex", booktabs = TRUE, caption = "ANOVA Table")
# Actual p-value = 0.00000000000000000000000000000000002955019, rounded to 0 in the Table
```
The structure of the dataset contains 2 columns, one includes all 6 scenarios, and the other shows every feeling score for the scenario. Conducting one-way ANOVA, we got a p-value < 0.005, as Table 1 shows. The statistical significance shows that there are difference in feelings across these 6 scenarios.\

## B.2 Conduct Pairwise t-tests
```{r paired t-tests, message=FALSE, warning=FALSE}
# Exclude the first column 'feelings_youalone' for the pairwise comparisons
scenarios <- names(feelings)[-1] 

# Initialize an empty list to store test results and a vector for p-values
pairwise_tests <- list()
p_values <- numeric(length(scenarios))

# Perform pairwise t-tests for all pairs where 'feelings_youalone' is one group
for (scenario in scenarios) {
  test_result <- t.test(feelings[[scenario]], 
                        feelings[['feelings_youalone']], paired = TRUE)
  pairwise_tests[[scenario]] <- test_result
  scenario_index <- which(names(feelings) == scenario)
  # Adjusting index as 'feelings_youalone' is at index 1
  p_values[scenario_index - 1] <- test_result$p.value 
}

# Adjust for multiple testing using Bonferroni correction for multiple comparison 
p_adjusted <- p.adjust(p_values, method = "bonferroni")

# Combine results into a data frame
results <- data.frame(
  Scenario = scenarios,
  P_Value = p_values,
  P_Adjusted = p_adjusted,
  options(scipen=9999))
print(results)
```
Given the ANOVA has a significant result, we next proceed to pairwise t-tests to compare variable "feelings_youalone" to each of the other 5 scenarios. As results shows: 
- "feelings_youalone" is significantly different from "feelings_bothyoufirst" (p < 0.01), "feelings_themalone" (p < 0.01), and "feelings_boththemfirst" (p < 0.01) after adjusting for multiple comparisons.
- Therefore, the significance suggests that feelings experienced when participants get no  apology is significantly different compared to the feelings experienced when they receive an apology, whether or not they being the first to apologize.

## C. The proportion of people choosing different options
```{r binary, message=FALSE, warning=FALSE}
outcome_table <- df %>%
  mutate(outcome_binary1 = case_when(
    outcome_binary1 == "I apologize first, then ${e://Field/initials} apologizes." ~ 
      "I apologize first, then others apologize.",
    outcome_binary1 == "Neither I nor ${e://Field/initials} apologizes." ~ 
      "Neither I nor others apologize.",
    TRUE ~ as.character(outcome_binary1))) %>%
  count(outcome_binary1) %>%
  mutate(proportion = n / sum(n))

# Plot the distribution of responses
ggplot(data = as.data.frame(outcome_table), aes(x = outcome_binary1, 
                                                y = proportion, 
                                                fill = outcome_binary1)) +
  geom_bar(stat = "identity") +
  labs(title = "Distribution of Responses for 'outcome binary1'", 
       x = "Response Option", 
       y = "Proportion") +
  scale_fill_manual(values = c(
    "I apologize first, then others apologize." = "darkgreen", 
    "Neither I nor others apologize." = "darkred")) +  
  geom_text(aes(label = sprintf("%.2f", proportion)),  
            vjust = -0.5, 
            size = 3) +
  theme(axis.text.x = element_text(angle = 20, hjust = 1, size = 6),
    panel.background = element_rect(fill = "white"),   
    panel.grid.major.y = element_line(color = "lightgray", linetype = 1, size = .5),
    panel.grid.major.x = element_line(color = "lightgray", linetype = 1, size = .5),
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank())


```
	
Again, think about your argument/conflict, and imagine no one has yet apologized. 